### 2017-12-17 Kokkos_PG_Views.tex
Reformat lines overrunning margins
Show how to make a View of Views; explain why this is generally unwise

\subsection{Constructing a View}

@@

\subsection{What types of data may a View contain?}

C++ lets users construct data types that may ``look like'' numbers in
terms of syntax, but do arbitrarily complicated things inside.  Some
of those things may not be thread safe, like unprotected updates to
global state.  Others may perform badly in parallel, like fine-grained
dynamic memory allocation.  Therefore it is strongly advised to use 
only simple data types inside Kokkos Views.
Users may always construct a View whose entries are 
\begin{itemize}
\item built-in data types (``plain old data''), like \texttt{int} or
  \texttt{double}, or
\item structs of built-in data types.
\end{itemize}

While it is in principal possible to have Kokkos Views of arbitrary 
objects there are a number of restrictions. The \lstinline|T| type 
must have a default constructor and destructor. For portability reasons
\lstinline|T| is not allowed to have virtual functions and one is not 
allowed to call functions with dynamic memory allocations inside of 
kernel code. Furthermore assignment operators as well as default constructor
and deconstructor must be marked with \lstinline|KOKKOS_[INLINE_]FUNCTION|. 
Keep in mind that the semantics of the resulting View are a combination of the 
Views 'view' semantics and the behaviour of the element type.

@@

\begin{enumerate}
\item Memory Spaces must match
\subitem \lstinline{ Kokkos::View<int*> -> Kokkos::View<int*,HostSpace> // ok if default memory space is HostSpace}
\item Data Type and Rank has to Match
\subitem \lstinline{int*  -> int*          //ok}
\subitem \lstinline{int*  -> const int*    //ok}
\subitem \lstinline{const int*  -> int*    //not ok, const violation}
\subitem \lstinline{int** -> int*          //not ok, rank mismatch}
\subitem \lstinline{int*[3] -> int**       //ok}
\subitem \lstinline{int** -> int*[3]       //ok if runtime dimension check matches}
\subitem \lstinline{int*  -> long*         //not ok, type mismatch}
\item Layouts must be compatible
\subitem \lstinline{LayoutRight -> LayoutRight //ok}
\subitem \lstinline{LayoutLeft -> LayoutRight  //not ok}
\subitem \lstinline{LayoutLeft -> LayoutSride  //ok}
\subitem \lstinline{LayoutStride -> LayoutLeft //ok if runtime dimensions allow assignment}
\item Memory Traits 
\end{enumerate}

>>>>

\subsection{Constructing a View}
\label{SS:View:CreateUse:Constructing}

@@

\subsection{What types of data may a View contain?}
\label{SS:View:CreateUse:ValidValueType}

C++ lets users construct data types that may ``look like'' numbers in
terms of syntax, but do arbitrarily complicated things inside.  Some
of those things may not be thread safe, like unprotected updates to
global state.  Others may perform badly in parallel, like fine-grained
dynamic memory allocation.  Therefore, we strongly advise that users
only store simple data types inside Kokkos Views.

Users may always construct a View whose entries are either
\begin{itemize}
\item built-in data types (``plain old data''), like \texttt{int} or
  \texttt{double}, or
\item structs of built-in data types.
\end{itemize}

It is possible to make Kokkos Views of other kinds of objects.
However, Kokkos imposes restrictions on the set of types \lstinline|T|
for which one can construct a \lstinline|View<T*>|.  For example:
\begin{enumerate}
\item \lstinline|T| must have a default constructor and destructor. 
\item \lstinline|T| must not have virtual methods.
\item \lstinline|T|'s default constructor and destructor must not
  allocate data, and must be thread safe.
\item \lstinline|T|'s assignment operators as well as its default
  constructor and deconstructor must be marked with the
  \lstinline|KOKKOS_INLINE_FUNCTION| or \lstinline|KOKKOS_FUNCTION|
  macro.
\end{enumerate}

All restrictions but the first come from the requirement that
\lstinline|View<T*>| work with every execution and memory space,
including those that use CUDA.  The constructor of
\lstinline|View<T*>| does not just allocate memory; it also
initializes the allocation by iterating over it using a
\lstinline|Kokkos::parallel_for|, and invoking \lstinline|T|'s default
constructor for each entry.  If the View's execution space is
\lstinline|Cuda|, then that means \lstinline|T|'s default constructor
needs to be correct to call on device.

\subsection{Can I make a View of Views?}
\label{SS:View:CreateUse:ViewOfViews}

A ``View of Views'' is a special case of View, where the type of each
entry is itself a View.  It is possible to make this, but before you
try, please see below.

\subsubsection{You probably don't want this}

If you really just want a multidimensional array, please don't do
this.  Instead, see Section \ref{SS:View:CreateUse:Constructing} for
the correct syntax.

If you want to represent an array of arrays, and the inner arrays have
fixed length or a fixed upper bound on length, consider instead using
a ``compressed sparse row'' data structure.  Kokkos' Containers
subpackage as a \lstinline|StaticCrsGraph| class that you may use for
this purpose.

If you want a hash table, Kokkos' Containers subpackage has an
\lstinline|UnorderedMap| class that you may use for this purpose.

One reason you might \emph{actually} want a View of Views is because
you need a representation of a ``ragged'' array of arrays -- where the
inner arrays have widely varying length -- and you need to be able to
reallocate the inner arrays dynamically.

You might also want a View of some class that itself contains Views.
If you want this, first think about how to reorganize your data
structures for better efficiency.

\subsubsection{I really want a View of Views; what do I do?}

Section \ref{SS:View:CreateUse:ValidValueType} above explains how the
outer View's constructor works.  The outer View's constructor does not
just allocate memory; it also initializes the allocation by iterating
over it using a \lstinline|Kokkos::parallel_for|, and invoking the
entry type's default constructor for each entry.  If the View's
execution space is \lstinline|Cuda|, then that means the entry type's
default constructor needs to be correct to call on device.  That is a
problem, because the entry type in this case is itself View.  View's
constructor wants to allocate memory, and thus does not work on
device.  Kokkos parallel regions generally forbid memory allocation.

One work-around is to invoke all inner Views' constructors on host.
For example, one might allocate a host mirror version of the outer
View, do the inner View allocations on host, then deep copy the outer
View to device.  Alternately, one might use \lstinline!CudaUVMSpace!
for the outer View, do the inner View allocations on host, then issue
a \lstinline!Kokkos::fence! before using the outer View on device.
The fence ensures that no device kernel will attempt to use the outer
View before the host initialization has finished.

A different approach defers constructing the inner Views until a
separate Kokkos parallel region.  One first constructs the outer View
using the \lstinline!WithoutInitializing! option, like this:
\begin{lstlisting}
using Kokkos::View;
using Kokkos::view_alloc;
using Kokkos::WithoutInitializing;

// Need an std::string here, because the compiler gets confused
// if you pass view_alloc a char* as its first argument.
const std::string label ("v_outer");
View<View<int*>> v_outer (view_alloc (label, WithoutInitializing));
\end{lstlisting}
This ensures that the inner Views' constructors don't get called on
device.  However, that also means that the inner Views are not
correctly constructed; they are not yet safe to use!  Thus, one must
then use ``placement new'' in a Kokkos parallel region, in order to
construct the inner Views.  This approach works best when the inner
Views share a common memory pool for all their allocations.

Here is a summary of different ways to make a View of Views:
\begin{enumerate}
\item Allocate a host mirror version of the outer View, do the inner
  View allocations on host, then deep copy the outer View to device.
\item Use \lstinline!CudaUVMSpace! for the outer View, do the inner
  View allocations on host, then fence before using the outer View on
  device.
\item Allocate the outer View using the
  \lstinline!WithoutInitializing! option, then use ``placement new''
  in a Kokkos parallel region to construct the inner Views.
\end{enumerate}

@@

\begin{enumerate}
\item Memory Spaces must match
  \subitem \lstinline{ Kokkos::View<int*> -> Kokkos::View<int*,HostSpace>}
  \subitem \lstinline{                      //ok if default memory space is HostSpace}
\item Data Type and Rank has to Match
\subitem \lstinline{int*  -> int*          //ok}
\subitem \lstinline{int*  -> const int*    //ok}
\subitem \lstinline{const int*  -> int*    //not ok, const violation}
\subitem \lstinline{int** -> int*          //not ok, rank mismatch}
\subitem \lstinline{int*[3] -> int**       //ok}
\subitem \lstinline{int** -> int*[3]       //ok if runtime dimension check matches}
\subitem \lstinline{int*  -> long*         //not ok, type mismatch}
\item Layouts must be compatible
\subitem \lstinline{LayoutRight -> LayoutRight //ok}
\subitem \lstinline{LayoutLeft -> LayoutRight  //not ok}
\subitem \lstinline{LayoutLeft -> LayoutSride  //ok}
\subitem \lstinline{LayoutStride -> LayoutLeft //ok if runtime dimensions allow assignment}
\item Memory Traits 
\end{enumerate}
